Globally-centered autocovariances in MCMC

### 2020-08-27 Paper/main.tex
Layer2 and a bit of Layer3
I will give it one last read today and then give it back to you.

% This gives us a full line break when we write a new paragraph
% \usepackage[mathscr]{euscript}
% \usepackage{tikz}
% \newcommand*\circled[1]{\tikz[baseline=(char.base)]{
%             \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
% \newcommand{\indep}{\perp \!\!\! \perp}

@@

\onehalfspacing
\begin{abstract}
Autocovariances are the fundamental quantity in many features of Markov chain Monte Carlo (MCMC) simulations with autcorrelation function (ACF) plots being often used as a visual tool to ascertain the performance of a Markov chain. Unfortunately, for slow mixing Markov chains, the empirical autocovariance can highly underestimate the truth. For multiple chain MCMC sampling,  we propose a globally-centered estimate of the autocovariance that pools information from all Markov chains. We show the impact of these improved estimators in three aspects: (1) acf plots, (2) estimates of the Monte Carlo asymptotic covariance matrix, and (3) estimates of the effective sample size. 
\end{abstract}

@@

A key quantity that drives MCMC output analysis is the autocovariance function (ACvF). From their use in autocorrelation (ACF) plots to assessing Monte Carlo variability of estimators, to determining when to stop MCMC simulation, autocovariances drive many visual and quantitative inferences from MCMC output. However, estimators of ACvF are only available for single-chain implementations, with multiple-chain version obtained by naive averaging. As we will demonstrate, this defeats the purpose of running parallel Markov chains from dispersed starting values.

Let $F$ be the target distribution with mean $\mu$ defined on a space $\X \subseteq \mathbb{R}^d$ equipped with a countably generated $\sigma$-field $\mathcal{B}(\X)$. For $s = 1, \dots, m$, let $\{X_{st}\}_{t\geq1}$ be the $s${th} Harris ergodic $F$-invariant Markov chain \citep[see][for definitions]{meyn:twee:2009} employed to learn characteristics about $F$. The process is covariance stationary so that
the ACvF at lag $k$ is defined as 
%
\[
    \Gamma(k) = \Cov_F(X_{s1}, X_{s\,1+k})= \mathbb{E}_F \left[(X_{s1} - \mu)(X_{s\,1+k} - \mu)^T \right]\,.
\]
Estimating $\Gamma(k)$ is critical to assessing the quality of the sampler and the reliability of Monte Carlo estimators. Let $\bar{X}_s = n^{-1} \sum_{t=1}^{n} X_{st}$ denote the Monte Carlo estimator of $\mu$ from the $s$th chain. The standard estimator for $\Gamma(k)$ is the sample autocovariance matrix at lag $k \geq 0$:
%
\begin{equation} \label{eq:empirical_ACvF}
    \hat{\Gamma}_s(k) = \dfrac{1}{n}\sum_{t=1}^{n-k} \left(X_{st} - \bar{X}_s \right) \left(X_{s\, t + k} - \bar{X}_s \right)^T\,,
\end{equation}
and for $k < 0$, $\hat{\Gamma}_s(k) = \hat{\Gamma}_s(-k)^T$. For single-chain MCMC runs, the estimator $\hat{\Gamma}_s(k)$ is used to construct ACF plots, to estimate the long-run variance of Monte Carlo estimators \citep{hannan:1970,dame:1991}, and to estimate effective sample size (ESS) \citep{kass:carlin:gelman:neal:1998,gong:fleg:2016,vats:fleg:jon:2019}. However, there is no unified approach to constructing estimators of $\Gamma(k)$ for multiple-chain implementations. Specifically, parallel MCMC chains are often spread across the state space so as to adequately capture high density areas. For slow mixing chains, the chains take time to traverse the space so that $\bar{X}_s$  are all vastly different. Consequently, $\hat{\Gamma}_s$ typically underestimates $\Gamma$ leading to a false sense of security about the quality of process.

@@

@@

\begin{ass}
\label{ass:lag_window}
The lag window $w(x)$ is continuous at $0$ at all but a finite number of points, is a bounded and even function with $w(0)=1$, \; $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $\sum_{k = -\infty}^{\infty} w(k/b_n) < \infty$.
\end{ass}

@@

The SV estimator, despite having good statistical properties, poses limitations due to slow computation. From \eqref{eq:empirical_ACvF} and \eqref{eq:sve}, the complexity of the SV estimator is $\mathcal{O}(n^2 p^2)$ in general, and for the Bartlett lag window, it is $\mathcal{O}(b_n n p^2)$. For slow mixing Markov chains, $n$ and $b_n$ are large, limiting the use of SV estimators.  To overcome this, we adapt the  fast Fourier transform based algorithm of \cite{heberle2017fast}. Suppose $w_k = w(k/b_n)$ and let $T(w)$ be the $n \times n$ Toeplitz matrix with the first column being $(1 ~ w_1 ~ w_2 ~ \dots, ~ w_{n-1})^T$. Notice an alternate formulation of $\hat{\Sigma}_s$
%
\begin{equation} \label{eq:kyriakoulis}
    \hat{\Sigma}_s = \dfrac{1}{n}A_s^T T(w) A_s, \qquad \textrm{ where } A_s = \begin{pmatrix}
    Y_{s1} - \bar{Y}_s  & \dots & Y_{sn} - \bar{Y}_s
\end{pmatrix}^T
\end{equation}
 \cite{heberle2017fast} compute the $n \times 1$ vector $T(w)A_s$  using an algorithm based on fast Fourier transforms. Let $w^* = (1 ~ w_1 ~ w_2 ~ \dots, ~ w_{n-1}, ~0, ~w_{n-1}, \dots, w_1)^T$ and set $C(w^*)$ to be a symmetric circulant matrix such that the matrix truncation $C(w^*)_{1:n, 1:n} = T(w)$.  With inputs $C(w^*)$ and $A_s$, Algorithm~\ref{algo:herberle} produces $\hat{\Sigma}_s$ exactly. For more details, see \cite{heberle2017fast}.
 % Let $C(w^*) = V\Lambda V^*$ be the spectral decomposition of $C(w^*)$, then $\Lambda$ is determined by a discrete Fourier transform of the first column of $C(w^*)$. 
% $A_s$ is extrapolated into a $2n \times p$ matrix $A_s^*$ by appending $n$ rows of $0$ vector to write Equation~\ref{eq:kyriakoulis} in the following form
% \[
%     \hat{\Sigma}_s = \dfrac{1}{n} A_s^T T(w) A_s = \dfrac{1}{n} A_s^T (C(w^*) A_s^*)_{1:n, :} = \dfrac{1}{n} A_s^T (V \Lambda V^* A_s^*)_{1:n, :}).
% \]
% Here $M_{1:r, \, 1:c}$ denotes the matrix truncation up to first $r$ rows and $c$ columns.

@@

In this section we consider three different target distributions and sample $m$ parallel Markov chains using MCMC methods to experimentally analyse the performance of our globally-centered estimators. We will make the following three comparisons - (1) A-ACF vs G-ACF through ACrF plots; (2) A-SVE vs G-SVE; (3) $\widehat{\textrm{ESS}}_G$ vs $\widehat{\textrm{ESS}}_A$. The quality of estimation of $\Sigma$ is studied by coverage probabilities of a $95 \%$ Wald  confidence interval in cases where the true mean $\mu$ is known. The convergence of local and global estimators of $\Sigma$ and $\widehat{\textrm{ESS}}$ as $n$ increases is studied through two types of running plots (1) logarithm of Frobenius norm of estimated $\Sigma$, and (2) logarithm of estimated ESS/$mn$. In all the examples we are interested in estimating the mean of the stationary distribution $\mu$, so $g$ is the identity function. The batch size $b_n$ is obtained from the function \texttt{batchSize} from \texttt{R} package \texttt{mcmcse}.

@@

\begin{lemma}
\label{lemma: brownian}
(\cite{csorgo2014strong}). Suppose the conditions of Theorem~\ref{thm:kuelbs} hold, then for all $\epsilon > 0$ and for almost all sample paths, there exists $n_{0}\left(\epsilon\right)$ such that $\forall n\geq n_{0}$ and $\forall i = 1, ..., p$

\[
\sup_{0\leq t \leq n-b_n}\sup_{0 \leq s \leq b_n} \left| B^{\left(i\right)}\left(t+s\right) - B^{\left(i\right)}\left(t\right) \right| < \left(1+ \epsilon\right)\left(2b_n\left(\log\dfrac{n}{b_n} + \log\; \log\; n\right)\right)^{1/2} ,
\]
%
\[
\sup_{0 \leq s \leq b_n} \left|B^{\left(i\right)}\left(n\right) - B^{\left(i\right)}\left(n - s\right)\right| < \left(1+ \epsilon\right)\left(2b_n\left(\log\dfrac{n}{b_n} + \log\;\log\;n\right)\right)^{1/2} , \;and
\]
%
\[
\left|B^{\left(i\right)}\left(n\right)\right| < \left(1+\epsilon\right)\sqrt{2n\;\log \log n} \; . 
\]
\end{lemma}

@@

@@

\begin{align*}
\hat{\Gamma}_{G,s}(k) & = \dfrac{1}{n}\sum_{t=1}^{n-|k|} \left(X_{st} - \bar{\bar{X}} \right) \left(X_{s(t+k)} - \bar{\bar{X}} \right)^T \\
    &= \left[\dfrac{1}{n} \sum_{t=1}^{n-|k|} \left(X_{st} - \bar{X}_s \right) \left(X_{s(t+k)} - \bar{X}_s \right)^T \right] + \left[\dfrac{1}{n} \sum_{t=1}^{|k|} \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - X_{st} \right)^T\right] \\ 
    & \quad +  \left[\dfrac{1}{n} \sum_{t=n-|k|+1}^{n}  \left( \bar{X}_s - X_{st} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right] + \left[\dfrac{n-|k|}{n} \left(\bar{X}_s - \bar{\bar{X}}\right)   \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right]\\
    &= \hat{\Gamma}_s(k) - \dfrac{1}{n} \sum_{t=1}^{|k|}A_{st}^T - \dfrac{1}{n}\sum_{t=n-|k|+1}^{n}A_{st} + \dfrac{n-|k|}{n}  \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\,, \numberthis \label{eq:acf_breakdown}
\end{align*}
%
where $A_{st} = (X_{st}-\bar{X}_s)(\bar{X}_s - \bar{\bar{X}})^T$. Under the assumption of stationarity, we will study the expectations of the each of the above terms. Without loss of generality, consider $A_{11}$,
\begin{align*}
    &\E \left[A_{11} \right]\\
     &= \mathbb{E} \left[ \left(X_{11} - \bar{X}_1 \right) \left(\bar{X}_1 - \bar{\bar{X}} \right)^T \right]\\
     % &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E}\left[X_{11}\bar{\bar{X}}^T \right] + \mathbb{E} \left[\bar{X}_1\bar{\bar{X}}^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{1}{m}\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{m-1}{m}\mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \dfrac{1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right] + \dfrac{m-1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \dfrac{m-1}{m}\left(\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\right)\\
    &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{t=1}^{n}\mathbb{E} \left[X_{11}X_{1t}^T \right] - \mathbb{E}\left[X_{11}\right] \mathbb{E}\left[\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\right]\mathbb{E}\left[\bar{X}_2^T\right] - \text{Var}\left[\bar{X}_1 \right] - \mathbb{E}\left[\bar{X}_1 \right]\mathbb{E}\left[\bar{X}_1^T \right]\right)\\
    % &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{k=0}^{n-1}\Gamma(k) + \mu \mu^T - \mu \mu^T + \mu \mu^T - \text{Var}\left[\bar{X}_1 \right] - \mu \mu^T\right)\\
    &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - n\text{Var}\left[\bar{X}_1 \right]\right)\,. \numberthis \label{eq:acf_2}
\end{align*}
%
Similarly,
%    
\begin{align}
\label{eq:acf_3}
    \mathbb{E} \left[ A_{11}^T \right] &= \mathbb{E}\left[ A_{11}\right]^T = \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k)^T - n\text{Var}\left[\bar{X}_1 \right] \right)\,.
    % &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - \Sigma \right)\\
\end{align}

Further,
\begin{align*}
\mathbb{E}\left[ \left(\bar{X}_{1} - \bar{\bar{X}} \right)  \left(\bar{X}_{1} - \bar{\bar{X}} \right)^{T}\right] &= \mathbb{E}\left[ \bar{X}_{1}\bar{X}_{1}^{T} - \bar{X}_{1}\bar{\bar{X}}^{T} - \bar{\bar{X}} \bar{X}_{1}^{T} + \bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
% &= \mathbb{E}\left[ \bar{X}_{1}\bar{X}_{1}^{T} - \bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
% &= \left(\mathbb{E}\left[\bar{X}_{1}\bar{X}_{1}^{T}\right] - \mathbb{E}\left[\bar{\bar{X}}\bar{\bar{X}}^{T}\right]\right)\\
&= \left(\Var(\bar{X}_{1}) + \mu\mu^{T} - \Var(\bar{\bar{X}}) - \mu\mu^{T}\right)\\
&= \dfrac{m-1}{m}\Var(\bar{X}_1)\,. \numberthis \label{eq:acf_4}
\end{align*}
%
Additionally, the locally-centered autocovariance exhibits the following expectation (from \cite{priestley1981spectral})
 \begin{equation} \label{eq:priestly}
     \mathbb{E}[\hat{\Gamma}(k)] = \left(1- \dfrac{\abs{k}}{n}\right)\left(\Gamma(k) - \Var{(\bar{X})}
 \right)\,.
 \end{equation}
%
As a consequence, if $\Var(\bar{X})$ is finite, then $\Var(\bar{X}) \to 0 \textrm{ as } n \to \infty$.
Using \eqref{eq:priestly}, \eqref{eq:acf_2}, \eqref{eq:acf_3}, and \eqref{eq:acf_4} in \eqref{eq:acf_breakdown},
\begin{align*}
    & \E \left[\hat{\Gamma}_{G,s}(k) \right] \\
    &= \mathbb{E}\left[\hat{\Gamma}_{s}(k)\right] - \dfrac{1}{n} \left(\sum\limits_{t=1}^{|k|}\mathbb{E}[A_{1t}^T] + \sum\limits_{t=n-|k|+1}^{n}\mathbb{E}[A_{1t}]\right) + \left(1- \dfrac{|k|}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)\\
    &= \mathbb{E}\left[\hat{\Gamma}_{s}(k)\right] - \dfrac{|k|}{n}\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h) + \dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)^T - 2 \Var(\bar{X}_1)\right) + \left(1- \dfrac{|k|}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)\\
    &= \left(1- \dfrac{|k|}{n}\right)\Gamma(k) - \dfrac{|k|}{n}\left[\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)^T + \dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)\right) - \left(2-\dfrac{1}{m}\right) \Var(\bar{X}_1)\right] - \dfrac{\Var(\bar{X}_1)}{m}\,.
\end{align*}
%

@@

\left|\hat{\Sigma}_{G}^{ij} - \tilde{\Sigma}^{ij}  \right| = |M_1 + M_2| \leq |M_1| + |M_2|\,.

@@

  \|\bar{Y}_s - \mu_g\|_{\infty} & \leq \|\bar{Y}_s - \mu_g\| = \dfrac{1}{mn}\left\|\sum_{t=1}^{n}Y_{st} - n\mu_g\right\| \\

@@

|M_1 + M_2  & \leq |M_1| + |M_2|

>>>>

@@

\onehalfspacing
\begin{abstract}
Autocovariances are the fundamental quantity in many features of Markov chain Monte Carlo (MCMC) simulations with autcorrelation function (ACF) plots being often used as a visual tool to ascertain the performance of a Markov chain. Unfortunately, for slow mixing Markov chains, the empirical autocovariance can highly underestimate the truth. For multiple chain MCMC sampling,  we propose a globally-centered estimator of autocovariance function (G-ACvF) that pools information from all Markov chains. G-ACvF is also shown to correct the negative bias of single chain empirical autocovariance. We show the impact of this improved estimator in three aspects: (1) ACF plots, (2) estimates of the Monte Carlo asymptotic covariance matrix, and (3) estimates of the effective sample size. We have established the strong consistency of our improved asymptotic variance estimator, and obtained its large sample bias and variance.
\end{abstract}

@@

A key quantity that drives MCMC output analysis is the autocovariance function (ACvF). From their use in autocorrelation (ACF) plots to assessing Monte Carlo variability of estimators, to determining when to stop MCMC simulation, autocovariances drive many visual and quantitative inferences from MCMC output. However, estimators of ACvF are only available for single-chain implementations, with parallel-chain version obtained by naive averaging. As we will demonstrate, this defeats the purpose of running parallel Markov chains from dispersed starting values.

Let $F$ be the target distribution with mean $\mu$ defined on a space $\X \subseteq \mathbb{R}^d$ equipped with a countably generated $\sigma$-field $\mathcal{B}(\X)$. For $s = 1, \dots, m$, let $\{X_{st}\}_{t\geq1}$ be the $s${th} Harris ergodic $F$-invariant Markov chain \citep[see][for definitions]{meyn:twee:2009} employed to learn characteristics about $F$. The process is covariance stationary so that
the ACvF at lag $k$ is defined as 
%
\[
    \Gamma(k) = \Cov_F(X_{s1}, X_{s\,1+k})= \mathbb{E}_F \left[(X_{s1} - \mu)(X_{s\,1+k} - \mu)^T \right]\,.
\]
Estimating $\Gamma(k)$ is critical to assessing the quality of the sampler and the reliability of Monte Carlo estimators. Let $\bar{X}_s = n^{-1} \sum_{t=1}^{n} X_{st}$ denote the Monte Carlo estimator of $\mu$ from the $s$th chain. The standard estimator for $\Gamma(k)$ is the sample autocovariance matrix at lag $k \geq 0$:
%
\begin{equation} \label{eq:empirical_ACvF}
    \hat{\Gamma}_s(k) = \dfrac{1}{n}\sum_{t=1}^{n-k} \left(X_{st} - \bar{X}_s \right) \left(X_{s\, t + k} - \bar{X}_s \right)^T\,,
\end{equation}
and for $k < 0$, $\hat{\Gamma}_s(k) = \hat{\Gamma}_s(-k)^T$. For single-chain MCMC runs, the estimator $\hat{\Gamma}_s(k)$ is used to construct ACF plots, to estimate the long-run variance of Monte Carlo estimators \citep{hannan:1970,dame:1991}, and to estimate effective sample size (ESS) \citep{kass:carlin:gelman:neal:1998,gong:fleg:2016,vats:fleg:jon:2019}. However, there is no unified approach to constructing estimators of $\Gamma(k)$ for parallel-chain implementations. Specifically, parallel MCMC chains are often spread across the state space so as to adequately capture high density areas. For slow mixing chains, the chains take time to traverse the space so that $\bar{X}_s$  are all vastly different. Consequently, $\hat{\Gamma}_s$ typically underestimates $\Gamma$ leading to a false sense of security about the quality of process.

@@

{\color{orange} I don't think that the RHS below is correct. In the formulation below, $o(n^{-2})$ is be $o(n^{-1})$}

@@

\begin{ass}
\label{ass:lag_window}
The lag window $w(x)$ is continuous at $0$ at all but a finite number of points, is a bounded and even function with $w(0)=1$, \; $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $\sum_{k = -\infty}^{\infty} \abs{w(k/b_n)} < \infty$.
\end{ass}

@@

The SV estimator, despite having good statistical properties, poses limitations due to slow computation. From \eqref{eq:empirical_ACvF} and \eqref{eq:sve}, the complexity of the SV estimator is $\mathcal{O}(n^2 p^2)$ in general, and for the Bartlett lag window, it is $\mathcal{O}(b_n n p^2)$. For slow mixing Markov chains, $n$ and $b_n$ are large, limiting the use of SV estimators.  To overcome this, we adapt the  fast Fourier transform based algorithm of \cite{heberle2017fast}. Suppose $w_k = w(k/b_n)$ and let $T(w)$ be the $n \times n$ Toeplitz matrix with the first column being $(1 ~ w_1 ~ w_2 ~ \dots, ~ w_{n-1})^T$. Notice an alternate formulation of $\hat{\Sigma}_s$
%
\begin{equation} \label{eq:kyriakoulis}
    \hat{\Sigma}_s = \dfrac{1}{n}A_s^T T(w) A_s, \qquad \textrm{ where } A_s = \begin{pmatrix}
    Y_{s1} - \bar{Y}_s  & \dots & Y_{sn} - \bar{Y}_s
\end{pmatrix}^T
\end{equation}
 \cite{heberle2017fast} compute the $n \times 1$ vector $T(w)A_s$  using an algorithm based on fast Fourier transforms. Let $w^* = (1 ~ w_1 ~ w_2 ~ \dots, ~ w_{n-1}, ~0, ~w_{n-1}, \dots, w_1)^T$ and set $C(w^*)$ to be a symmetric circulant matrix such that the matrix truncation $C(w^*)_{1:n, 1:n} = T(w)$. Let $M_{(j)}$ denote the $j$th column of matrix $M$ and $v^{(i)}$ denote the $i$th element of the vector $v$. With inputs $C(w^*)$ and $A_s$, Algorithm~\ref{algo:herberle} produces $\hat{\Sigma}_s$ exactly. For more details, see \cite{heberle2017fast}.
 % Let $C(w^*) = V\Lambda V^*$ be the spectral decomposition of $C(w^*)$, then $\Lambda$ is determined by a discrete Fourier transform of the first column of $C(w^*)$. 
% $A_s$ is extrapolated into a $2n \times p$ matrix $A_s^*$ by appending $n$ rows of $0$ vector to write Equation~\ref{eq:kyriakoulis} in the following form
% \[
%     \hat{\Sigma}_s = \dfrac{1}{n} A_s^T T(w) A_s = \dfrac{1}{n} A_s^T (C(w^*) A_s^*)_{1:n, :} = \dfrac{1}{n} A_s^T (V \Lambda V^* A_s^*)_{1:n, :}).
% \]
% Here $M_{1:r, \, 1:c}$ denotes the matrix truncation up to first $r$ rows and $c$ columns.

@@

In this section we consider three different target distributions and sample $m$ parallel Markov chains using MCMC methods to experimentally analyse the performance of our globally-centered estimators. We will make the following three comparisons - (1) A-ACF vs G-ACF through ACrF plots; (2) A-SVE vs G-SVE; (3) $\widehat{\textrm{ESS}}_G$ vs $\widehat{\textrm{ESS}}_A$. The quality of estimation of $\Sigma$ is studied by coverage probabilities of a $95 \%$ Wald  confidence interval in cases where the true mean $\mu$ is known. The convergence of local and global estimators of $\Sigma$ and $\widehat{\textrm{ESS}}$ as $n$ increases is studied through two types of running plots (1) logarithm of Frobenius norm of estimated $\Sigma$, and (2) logarithm of estimated ESS/$mn$. In all the examples we are interested in estimating the mean of the stationary distribution $\mu$, so $g$ is the identity function. The batch size $b_n$ is obtained from the function \texttt{batchSize} from \texttt{R} package \texttt{mcmcse} \citep{mcmcse}.

@@

\begin{lemma}
\label{lemma: brownian}
(\cite{csorgo1981strong}). Suppose the conditions of Theorem~\ref{thm:kuelbs} hold, then for all $\epsilon > 0$ and for almost all sample paths, there exists $n_{0}\left(\epsilon\right)$ such that $\forall n\geq n_{0}(\epsilon)$ and $\forall i = 1, ..., p$

\[
\sup_{0\leq t \leq n-b_n}\sup_{0 \leq s \leq b_n} \left| B^{\left(i\right)}\left(t+s\right) - B^{\left(i\right)}\left(t\right) \right| < \left(1+ \epsilon\right)\left(2b_n\left(\log\dfrac{n}{b_n} + \log\; \log\; n\right)\right)^{1/2} ,
\]
%
\[
\sup_{0 \leq s \leq b_n} \left|B^{\left(i\right)}\left(n\right) - B^{\left(i\right)}\left(n - s\right)\right| < \left(1+ \epsilon\right)\left(2b_n\left(\log\dfrac{n}{b_n} + \log\;\log\;n\right)\right)^{1/2} , \; \textrm{and}
\]
%
\[
\left|B^{\left(i\right)}\left(n\right)\right| < \left(1+\epsilon\right)\sqrt{2n\;\log \log n} \; . 
\]
\end{lemma}

@@

For the purpose of generalizing summations for positive and negative lags in the following proofs, we define the sets $I_k, J_{k1} \text{ and } J_{k2}$ as $ I_k := \{1, \dots, n-k\}, J_{k1} := \{n-k+1, \dots , n\}, J_{k2} := \{1, \dots, k\}$ for $k > 0$ and  $I_k := \{1-k, \dots, n\}, J_{k1} := \{1, \dots , -k\}, J_{k2} := \{n+k+1, \dots, n\}$ for $k < 0$. For $k=0$, $J_{k1}$ and $J_{k2}$ are empty sets and $I_k := \{1, \dots, n\}$. Notice that the empirical autocovariance estimator at lag-$k$ requires a summation over $I_k$.

@@

\begin{align*}
\hat{\Gamma}_{G,s}(k) & = \dfrac{1}{n}\sum_{t \in I_k} \left(X_{st} - \bar{\bar{X}} \right) \left(X_{s(t+k)} - \bar{\bar{X}} \right)^T \\
    &= \left[\dfrac{1}{n} \sum_{t \in I_k} \left(X_{st} - \bar{X}_s \right) \left(X_{s(t+k)} - \bar{X}_s \right)^T \right] + \left[\dfrac{1}{n} \sum_{t \in J_{k2}} \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - X_{st} \right)^T\right] \\ 
    & \quad +  \left[\dfrac{1}{n} \sum_{t \in J_{k1}}  \left( \bar{X}_s - X_{st} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right] + \left[\dfrac{n-\abs{k}}{n} \left(\bar{X}_s - \bar{\bar{X}}\right)   \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right]\\
    &= \hat{\Gamma}_s(k) - \dfrac{1}{n} \sum_{t \in J_{k2}}^{k}A_{st}^T - \dfrac{1}{n}\sum_{t \in J_{k1}}A_{st} + \dfrac{n-\abs{k}}{n}  \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\,, \numberthis \label{eq:acf_breakdown}
\end{align*}
%
where $A_{st} = (X_{st}-\bar{X}_s)(\bar{X}_s - \bar{\bar{X}})^T$. Under the assumption of stationarity, we will study the expectations of each of the above terms. Without loss of generality, consider $A_{11}$,
\begin{align*}
    &\E \left[A_{11} \right]\\
     &= \mathbb{E} \left[ \left(X_{11} - \bar{X}_1 \right) \left(\bar{X}_1 - \bar{\bar{X}} \right)^T \right]\\
     % &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E}\left[X_{11}\bar{\bar{X}}^T \right] + \mathbb{E} \left[\bar{X}_1\bar{\bar{X}}^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{1}{m}\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{m-1}{m}\mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \dfrac{1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right] + \dfrac{m-1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \dfrac{m-1}{m}\left(\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\right)\\
    &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{t=1}^{n}\mathbb{E} \left[X_{11}X_{1t}^T \right] - \mathbb{E}\left[X_{11}\right] \mathbb{E}\left[\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\right]\mathbb{E}\left[\bar{X}_2^T\right] - \text{Var}\left[\bar{X}_1 \right] - \mathbb{E}\left[\bar{X}_1 \right]\mathbb{E}\left[\bar{X}_1^T \right]\right)\\
    % &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{k=0}^{n-1}\Gamma(k) + \mu \mu^T - \mu \mu^T + \mu \mu^T - \text{Var}\left[\bar{X}_1 \right] - \mu \mu^T\right)\\
    &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - n\text{Var}\left[\bar{X}_1 \right]\right)\,. \numberthis \label{eq:acf_2}
\end{align*}
%
Similarly,
%    
\begin{align}
\label{eq:acf_3}
    \mathbb{E} \left[ A_{11}^T \right] &= \mathbb{E}\left[ A_{11}\right]^T = \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k)^T - n\text{Var}\left[\bar{X}_1 \right] \right)\,.
    % &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - \Sigma \right)\\
\end{align}

Further,
\begin{align*}
\mathbb{E}\left[ \left(\bar{X}_{1} - \bar{\bar{X}} \right)  \left(\bar{X}_{1} - \bar{\bar{X}} \right)^{T}\right] &= \mathbb{E}\left[ \bar{X}_{1}\bar{X}_{1}^{T} - \bar{X}_{1}\bar{\bar{X}}^{T} - \bar{\bar{X}} \bar{X}_{1}^{T} + \bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
% &= \mathbb{E}\left[ \bar{X}_{1}\bar{X}_{1}^{T} - \bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
% &= \left(\mathbb{E}\left[\bar{X}_{1}\bar{X}_{1}^{T}\right] - \mathbb{E}\left[\bar{\bar{X}}\bar{\bar{X}}^{T}\right]\right)\\
&= \left(\Var(\bar{X}_{1}) + \mu\mu^{T} - \Var(\bar{\bar{X}}) - \mu\mu^{T}\right)\\
&= \dfrac{m-1}{m}\Var(\bar{X}_1)\,. \numberthis \label{eq:acf_4}
\end{align*}
%
Additionally, the locally-centered autocovariance exhibits the following expectation (from \cite{priestley1981spectral})
 \begin{equation} \label{eq:priestly}
     \mathbb{E}[\hat{\Gamma}(k)] = \left(1- \dfrac{\abs{k}}{n}\right)\left(\Gamma(k) - \Var{(\bar{X})}
 \right)\,.
 \end{equation}
%
Using\eqref{eq:acf_2}, \eqref{eq:acf_3}, \eqref{eq:acf_4}, and  \eqref{eq:priestly} in \eqref{eq:acf_breakdown},
\begin{align*}
    & \E \left[\hat{\Gamma}_{G,s}(k) \right] \\
    &= \mathbb{E}\left[\hat{\Gamma}_{s}(k)\right] - \dfrac{1}{n} \left(\sum\limits_{t \in J_{k2}}\mathbb{E}[A_{1t}^T] + \sum\limits_{t \in J_{k1}}\mathbb{E}[A_{1t}]\right) + \left(1- \dfrac{\abs{k}}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)\\
    &= \mathbb{E}\left[\hat{\Gamma}_{s}(k)\right] - \dfrac{\abs{k}}{n}\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h) + \dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)^T - 2 \Var(\bar{X}_1)\right) + \left(1- \dfrac{\abs{k}}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)\\
    &= \left(1- \dfrac{\abs{k}}{n}\right)\Gamma(k) - \dfrac{\abs{k}}{n}\left[\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)^T + \dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)\right) - \left(2-\dfrac{1}{m}\right) \Var(\bar{X}_1)\right] - \dfrac{\Var(\bar{X}_1)}{m}\,.
\end{align*}
%

@@

\left|\hat{\Sigma}_{G}^{ij} - \tilde{\Sigma}_{A}^{ij}  \right| = |M_1 + M_2| \leq |M_1| + |M_2|\,.

@@

  \|\bar{Y}_s - \mu_g\|_{\infty} & \leq \|\bar{Y}_s - \mu_g\| = \dfrac{1}{n}\left\|\sum_{t=1}^{n}Y_{st} - n\mu_g\right\| \\

@@

|M_1 + M_2|  & \leq |M_1| + |M_2|
